Metabolites reflect variability introduced by mesoscale eddies in the North Pacific Subtropical Gyre

### 2024-10-22 manuscript/manuscript.Rmd
Adding version information and citations

The Waters Acquity UPLC was coupled to a Thermo Q Exactive HF hybrid Orbitrap high resolution mass spectrometer equipped with a heated electrospray ionization source (H-ESI). The H-ESI voltage was set to 3.3 kV and sheath gas, auxiliary gas, and sweep gas flow rates were set at 16, 3, and 1, respectively. The capillary and auxiliary gas heater temperatures were maintained at 320°C and 100°C, respectively. Full scan analyses were performed with polarity switching and a scan range of 60 to 900 *m/z* at a resolution of 60,000. The instrument was mass calibrated at 200 *m/z* before each run began and once again during the eddy transect sample set to ensure calibrations were always performed within 3-4 days. All data files were then converted to an open-source mzML format and vendor centroided via Proteowizard’s msConvert tool. All mzML files have been uploaded to Metabolomics Workbench under Project ID PR001738.

@@

Compounds for which we had an authentic standard (Supplemental Table 5) were manually integrated using Skyline and MSDIAL with the standard mixes used to ensure the correct peak was integrated by matching retention time. Compounds were removed from the analysis if they were visually assessed to have poor peak quality in the samples or peak areas in the samples similar to those in the methodological blanks. Skyline was used for both the eddy center samples and the HEE cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset. Raw peak areas were then normalized to their best-matched internal standard (Supplemental Table 4) per @Boysen2018 except that all compounds were normalized to their best match no matter how minimal the improvement. Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standard mixes and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.

@@

Samples were also processed via the untargeted workflow detailed in @Kumler2023. In short, XCMS [@R-xcms] was used for feature detection with CentWave peakpicking, Obiwarp alignment, and peak density correspondence/grouping. This was followed by manual inspection of the extracted ion chromatogram for each mass feature by an expert for quality with only those peaks classified as "Good" used in the analysis here. Features were then matched with and normalized to an internal standard (Supplemental Table 1) as described above.

@@

All statistics were run in R version 4.3.3. We used multivariate statistics provided by the vegan package [@R-vegan] to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses. Non-parametric tests were used when data were unlikely to obey parametric assumptions and permutational statistics were preferred wherever possible. Multiple comparison problems were controlled where necessary using a false-discovery rate correction [@Benjamini1995]. For the analyses involving the eddy transect samples, sea level anomaly was treated as a continuous variable with the exception of the PERMANOVA, for which SLA was categorized as cyclonic if less than -5 cm and anticyclonic if greater than 5 cm, with a third category for stations in between [@Anderson2001]. Both the MESO-SCOPE eddy center and HEE analyses treated SLA as a categorical variable.

>>>>

The Waters Acquity UPLC was coupled to a Thermo Q Exactive HF hybrid Orbitrap high resolution mass spectrometer equipped with a heated electrospray ionization source (H-ESI). The H-ESI voltage was set to 3.3 kV and sheath gas, auxiliary gas, and sweep gas flow rates were set at 16, 3, and 1, respectively. The capillary and auxiliary gas heater temperatures were maintained at 320°C and 100°C, respectively. Full scan analyses were performed with polarity switching and a scan range of 60 to 900 *m/z* at a resolution of 60,000. The instrument was mass calibrated at 200 *m/z* before each run began and once again during the eddy transect sample set to ensure calibrations were always performed within 3-4 days. All data files were then converted to an open-source mzML format and vendor centroided via Proteowizard’s msConvert tool (version 3.0.19297, @Chambers2012). All mzML files have been uploaded to Metabolomics Workbench under Project ID PR001738.

@@

Compounds for which we had an authentic standard (Supplemental Table 5) were manually integrated using Skyline (versions 4.1 and 21.2.0.470, @Adams2020) and MSDIAL (version 4.36, @Tsugawa2015) with the standard mixes used to ensure the correct peak was integrated by matching retention time. Compounds were removed from the analysis if they were visually assessed to have poor peak quality in the samples or peak areas in the samples similar to those in the methodological blanks. Skyline was used for both the eddy center samples and the HEE cruise data while MSDIAL was used for the eddy transect data due to the larger number of files in that dataset. Raw peak areas were then normalized to their best-matched internal standard (Supplemental Table 4) per @Boysen2018 except that all compounds were normalized to their best match no matter how minimal the improvement. Normalized peaks were then calibrated via a single-point standard curve as determined uniquely for each compound from the authentic standard mixes and used to convert normalized peak area into moles per µL injected and scaled to the amount of seawater filtered to provide a final estimate of environmental concentration for each compound in moles per liter of seawater filtered.

@@

Samples were also processed via the untargeted workflow detailed in @Kumler2023. In short, XCMS (version 3.10.0, @R-xcms) was used for feature detection with CentWave peakpicking, Obiwarp alignment, and peak density correspondence/grouping. This was followed by manual inspection of the extracted ion chromatogram for each mass feature by an expert for quality with only those peaks classified as "Good" used in the analysis here. Features were then matched with and normalized to an internal standard (Supplemental Table 1) as described above.

@@

All statistics were run in R version 4.4.0. We used multivariate statistics provided by the vegan package (version 2.6-6.1, @R-vegan) to assess the overall impact of various metrics across the metabolome followed by univariate statistics to investigate individually significant responses. Non-parametric tests were used when data were unlikely to obey parametric assumptions and permutational statistics were preferred wherever possible. Multiple comparison problems were controlled where necessary using a false-discovery rate correction [@Benjamini1995]. For the analyses involving the eddy transect samples, sea level anomaly was treated as a continuous variable with the exception of the PERMANOVA, for which SLA was categorized as cyclonic if less than -5 cm and anticyclonic if greater than 5 cm, with a third category for stations in between [@Anderson2001]. Both the MESO-SCOPE eddy center and HEE analyses treated SLA as a categorical variable.
